---
title: "Glados Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{glados}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(glados)
```

Named after the iconic AI character in the Portal video games, the glados package can be used for scientific computing in a variety of domains. Let's take a look at some of the applications in the current version of glados.

`euclidean`

The Euclidean distance measures the shortest possible distance between two points on a plane. Mathematically, it is defined as $d(p, q) = \sqrt{\sum_{i = 1}^{n}(q_{i} - p_{i})^{2}}$, where $p$ and $q$ represent two $n$-dimensional vectors (data points), and $p_{i}, q_{i}$ represent the $i$-th element of the respective vector.

**Applications:**

Suppose we want to compute the shortest distance from Ann Arbor to Glacier Bay, Alaska, disregarding any geographic or human-made barriers. Ann Arbor's [DD coordinates](https://en.wikipedia.org/wiki/Decimal_degrees#:~:text=Decimal%20degrees%20(DD)%20is%20a,as%20OpenStreetMap%2C%20and%20GPS%20devices.) are ([42.27756, -83.74088](https://latitude.to/map/us/united-states/cities/ann-arbor)), while Glacier Bay's are ([58.6999972, -136.1499994](https://latitude.to/articles-by-country/us/united-states/31627/glacier-bay-basin)).

```{r euclidean_geography}
ann_arbor <- c(42.27756, -83.74088) 
glacier_bay <- c(58.6999972, -136.1499994)
get_euclidean_distance(ann_arbor, glacier_bay)
```

Using our geographical knowledge, this means that Glacier Bay is roughly 55 degrees northwest of Ann Arbor!

**Baseline Comparisons:**

Although the built-in `dist()` function returns the same values, the glados version improves upon its efficiency, both computationally and with regards to UX design.

```{r dist}
# original function 
dist(rbind(ann_arbor, glacier_bay), 'euclidean')[1][1]
```

Notice the extra steps taken to compute the distance? For users with little experience in R, `rbind()` may not be intuitive, and it takes time to figure out. Furthermore, we can only isolate the numerical result by indexing [1][1]; otherwise, we would simply return a 1 x 1 matrix with the row and column names included.

```{r matrix}
dist(rbind(ann_arbor, glacier_bay), 'euclidean')
```

These two problems are solved with glados. The user is not required to concatenate the matrix, and the results are returned with a straightforward number. Due to this optimized code, `get_euclidean_distance` returns improved benchmarks.

```{r benchmark}
if(!require(bench)) install.packages("bench",repos = "http://cran.us.r-project.org")
library(bench) 
bench::mark(get_euclidean_distance(ann_arbor, glacier_bay),
            dist(rbind(ann_arbor, glacier_bay), 'euclidean')[1][1])
```

```{r complicated}
c <- c(0.104717,0.066009,-0.019987,-0.020995,-0.028313,-0.007809,0.012926,-0.007384,-0.043359,-0.018224,0,0.007343,-0.019177,0.005367,0.00665,-0.005701,0.019818,-0.00152,-0.003645,0.024512,0.010731,-0.00507,-0.002958,-0.009399,-0.002515)
d <- c(-0.626027,0.062963,0.018856,0.01292,-0.001846,0.010877,0.019271,-0.004384,-0.02127,0.002734,0.001786,-0.003597,-0.014717,0,-0.015336,0.0118,-0.002086,0.044721,-0.019986,0.005127,-0.005241,-0.001978,-0.00037,0.003012,-0.002036)
bench::mark(get_euclidean_distance(c, d),
            dist(rbind(c, d), 'euclidean')[1][1])
```

Improved statistics in glados include the following:

-   Minimum run time

-   Amount of garbage collections (GC)
